Enhancing Accelerometer-based Activity Recognition with Capacitive Proximity Sensing

### 2012-06-15 Paper/Paper.tex

@@

The 3D accelerometer sensor is being sampled at $100Hz$, resulting in 10ms equidistant measurements. For efficiency reasons, the sensor data is run-length encoded before being stored locally to the microSD card.

Since the main microcontroller on the sensor node has unoccupied pins, it can be extended with further sensors tailoring different application scenarios. For our scenario, we have added a capacitive proximity sensor, which is described in detail in the next section.

@@

The capacitive proximity sensor for the wristband performs measurements in loading mode. The sensing circuit schematic is shown in Figure \ref{}. It is based on a timer that controls the charging and discharging cycles of the virtual capacitor. The timer toggles from charge to discharge at the time when a threshold voltage at the capacitor is reached. This results in an astable operation with succeeding charge/discharge cycles. When the capacitance of the virtual capacitor increases, the charging time will also increase and vice-versa. Therefore, the capacitance is inversely proportional to the number of charging cycles in a given time span. In order to guard the sensor from measuring the capacitance to underlying objects, a shield is placed directly underneath the measuring electrode. The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is neglectable. Using this shielding method, the measured capacitance will only be slightly affected by the grounded underlying arm.

The proximity sensor board generates a clock signal with varying frequency depending on the proximity to body parts, objects and their characteristics. The HedgeHog therefore needs to measure the incoming frequency, which is implemented by counting the signal's edges over the gate time of approximately 9.5ms.

\begin{verbatim}
pseudocode:
  sample the accelerometer sensor
  start frequency counter
  go to sleep for 9.5 ms
  wake-up (timer interrupt)
  stop frequency counter
  read frequency count
\end{verbatim}

\section{Experiment}

@@

\begin{figure}
	\centering
		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/making-bread.pdf}
	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast. Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in ``unmittelbarer nähe''.}
	\label{fig:making-bread}
\end{figure}

>>>>

\begin{figure}[htbp]
	\centering
% 		\includegraphics[width=1.00\textwidth]{Images/modes.png}
	\caption{Capacitive proximity sensing modes \cite{Smith1996}}
	\label{fig:modes}
\end{figure}

@@

The 3D accelerometer sensor is being sampled at $100Hz$, resulting in 10ms equidistant measurements. For efficiency reasons, the sensor data is run-length encoded before being stored locally to the microSD card. The HedgeHog can be extended with further sensors tailoring different application scenarios. For our scenario, we have added a capacitive proximity sensor, which is described in detail in the next section.

@@

The capacitive proximity sensor for the wristband performs measurements in loading mode. The sensing circuit schematic is shown in Figure \ref{}. It is based on a timer that controls the charging and discharging cycles of the virtual capacitor that is built by the wristband electrode and the surrounding environment. The timer toggles from charge to discharge at the time when a threshold voltage at the capacitor is reached. This results in an astable operation with succeeding charge/discharge cycles. When the capacitance of the virtual capacitor increases, the charging time will also increase and vice-versa. Therefore, the capacitance is inversely proportional to the number of charging cycles in a given time span. In order to guard the sensor from measuring the capacitance to underlying objects, a shield is placed directly underneath the measuring electrode. The shield is driven with the same potential as the sensing electrode, such that the the capacitance between the two electrodes is negligible. Using this shielding method, the measured capacitance will only be slightly affected by the grounded underlying arm.

The proximity sensor board generates a clock signal with varying frequency depending on the charge and discharge cycles. The HedgeHog measures the resulting capacitance by counting the signal's edges over a gate time of approximately 9.5ms.

\begin{figure}[htbp]
	\centering
% 		\includegraphics[trim=1cm 1cm 1cm 1cm,clip,width=\textwidth]{Images/pseudocode.pdf}
	\caption{Measurement procedure}
	\label{fig:pseudocode}
\end{figure}
\section{Experiment}

@@

\begin{figure}
	\centering
		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/making-bread.pdf}
	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast. Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.}
	\label{fig:making-bread}
\end{figure}
